---
title: "Natural Disasters - Creating Viz"
author: "Jaslyn Miura"
format: pdf
date: 2026-03-08
---

**Goal**: The goal with this data is to explore how the types and amount of natural disasters that have occurred around the world have varied since 1960-2018.

**Questions of Interest:** 
What type natural disasters occur the most around the world?
Which countries experience the most natural disasters per year around the world?
How have the distribution of natural disasters changed over the years (type and count)? Could the changing trends identify major shifts, such as the rise of climate change?

```{r, warning=FALSE, message=FALSE}
# load necessary libraries
library(tidyverse)
library(dplyr)
library(paletteer)
library(lubridate)
library(showtext)
library(sysfonts)
library(stringr)
library(ggstream)
library(cowplot)
library(ggplot2)
library(rnaturalearth)
library(sf)
#install.packages("ggimage")
library(ggimage)
#install.packages("countrycode")
library(countrycode)
library(janitor)
library(gghighlight)
library(waffle)
```

## Loading Data
```{r, message=FALSE}
# read in natural disaster world data.
nat_dis <- read_csv(here::here("data", "pend-gdis-1960-2018-disasterlocations-csv", "pend-gdis-1960-2018-disasterlocations.csv")) %>%
  mutate(year = as.integer(year)) %>% 
  filter(disastertype %in% c("earthquake", "flood", "storm", "drought", "extreme temperature"))

# read in population world data.
pop <- read_csv(here::here("data", "world_population", "population.csv"), skip = 4, show_col_types = FALSE) %>% 
  select(where(~ !all(is.na(.))))

# read in displaced world data. (2008-2018)
displaced <- read_csv(here::here("data", 
                                 "internally-displaced-persons-from-disasters",
                                 "internally-displaced-persons-from-disasters.csv")) %>% 
  rename("displaced" = "Internally displaced persons, new displacement associated with disasters (number of cases)")
```

```{r}
# cleaning population data
pop_long <- pop %>% 
  pivot_longer(
    cols = `1960`:`2024`,
    names_to = "year",
    values_to = "population") %>% 
  mutate(year = as.integer(year)) %>% 
  clean_names() %>% 
  select(country_name, year, population, country_code) %>% 
  mutate(country_name = recode(country_name,
                               "United States of America" = "United States",
                               "Viet Nam" = "Vietnam"))
```


```{r}
# add special fonts
font_add_google(name = "Space Grotesk", family = "sans-serif") # axis details and numbers
font_add_google(name = "Bricolage Grotesque", family = "bricolage") # titles, captions

# creating a color palette
"grey_blue" <- "#8896A7" # bar color
"pale_blue" <- "#A1BBDD" # text color
"dark_grey" <- "#4D4D4D" # plot background details

# create a color palette for disasters
pal <- c("drought" = "#663317",
         "earthquake" = "#0B4A32",
         "extreme temperature" = "#783682",
         "flood" = "#75B9B0",
         "storm" = "#E89300")
```


### Line Graph
```{r, message=FALSE}
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                  Total Natural Disaster Trend (1960-2018)                ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

#.........................Summarize Data.........................

# Create the `yearly_total` dataframe.
yearly_total <- nat_dis %>% 
  # Get the total number of disasters per year.
  count(year, name = "total_disasters")

#............................Line Plot...........................

showtext.auto(enable = TRUE)

# Create a line graph.
line_trend <- ggplot(data = yearly_total, aes(x = year, y = total_disasters)) +
  geom_line() +
  geom_point(size = 0.4) +
  scale_x_continuous(limits = c(1960, 2018),
                     breaks = c(seq(1960, 2015, by = 20), 2018)) +
  scale_y_continuous(labels = scales::label_comma()) +
  labs(y = "Total Natural Disasters",
       ) +
  theme_minimal(base_size = 17) +
  # Theme adjustments.
  theme(
    axis.text = element_text(family = "sans-serif",
                             color = pale_blue),
    axis.text.y = element_text(size = rel(1.5), 
                               color = pale_blue),
    axis.text.x = element_text(size = rel(1.5),
                               color = pale_blue),
    axis.title.x = element_blank(),
    axis.title.y = element_text(family = "bricolage",
                                size = rel(1.8), 
                                margin = margin(t = 15,
                                                r = 12),
                                color = pale_blue),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(linetype = "dashed", 
                                      color = dark_grey),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background  = element_rect(fill = "transparent", color = NA))
showtext_opts(dpi = 300)

ggsave(
  filename = here::here("images", "line_trend.pdf"),
  plot = line_trend,
  device = cairo_pdf,
  width = 21,
  height = 5.2,
  units = "in")

ggsave(
  filename = here::here("images", "line_trend.png"),
  plot = line_trend,
  device = png,
  width = 21,
  height = 5.2,
  units = "in")

```

### Bar Graph with Flags
```{r}
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                     Total Natural Disasters by Country                   ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

#.........................Summarize Data.........................

# Create the `top_10` dataframe.
top10_df <- nat_dis %>% 
  # Get the yearly totals for each country.
  count(country, year, name = "disaster_count") %>% 
  # Calculate the total of all events between 1960-2018.
  group_by(country) %>% 
  summarize(country_sum = sum(disaster_count), 
            .groups = "drop") %>%  
  # Select the top 10 countries.
  slice_max(country_sum, n = 10) %>%    
  # Add flag image source as a new column.
  mutate(iso2 = countrycode(country, 
                            origin = "country.name", 
                            destination = "iso2c"),
         flag = paste0("https://flagcdn.com/w40/", tolower(iso2), ".png"),
         country = fct_reorder(country, country_sum))

#............................Bar Plot............................

showtext.auto(enable = TRUE)

# Create a bar chart.
top10 <- ggplot(top10_df, aes(x = country_sum, y = country)) +
  geom_col(fill = grey_blue,
           alpha = 0.5) +
  # Add the total value to the end of the bar.
  geom_text(aes(label = scales::comma(country_sum)),
            hjust = -0.4, 
            color = pale_blue,
            size = 7,
            family = "sans-serif") +
  # Add flag images to associated country.
  geom_image(aes(image = flag),
             x = max(top10_df$country_sum) * 0.03,
             size = 0.05,
             by = "width") +
  # Expand the x-axis for for flag and total value placement.
  scale_x_continuous(labels = scales::label_comma(),
                     expand = expansion(mult = c(0, 0.1))) +
  labs(x = "Number of Natural Disasters",) +
  theme_minimal(base_size = 17) +
  # Theme adjustments.
  theme(
    axis.text = element_text(family = "bricolage",
                             size = rel(1.5),
                             color = pale_blue),
    axis.text.y = element_text(size = rel(1),
                               color = pale_blue),
    axis.text.x = element_blank(),
    axis.title.x = element_text(family = "bricolage",
                                size = rel(2.5),
                                color = pale_blue),
    axis.title.y = element_blank(),
    panel.grid = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background  = element_rect(fill = "transparent", color = NA))

showtext_opts(dpi = 300)

ggsave(
  filename = here::here("images", "top10.pdf"),
  plot = top10,
  device = cairo_pdf,
  width = 14,
  height = 8,
  units = "in"
  )

ggsave(
  filename = here::here("images", "top10.png"),
  plot = top10,
  device = png,
  width = 14,
  height = 8,
  units = "in")


```


### Population Bar Chart 
```{r}
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                        Disaster Events by Population                     ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

#.........................Summarize Data.........................

# Create top 10 country list from referencing `top_10` bar plot.
top10_list <- c("United States", 
                "Philippines", 
                "India", 
                "China", 
                "Japan", 
                "Thailand", 
                "Indonesia", 
                "Bangladesh", 
                "Vietnam", 
                "Peru")

# Create `pop_avg_top10` dataframe.
pop_avg_top10 <- pop_long %>% 
  # Filter data by the designated time frame and countries of interest.
  filter(year >= 1960, year <= 2018) %>% 
  filter(country_name %in% top10_list) %>% 
  # Calculate the average population for each country.
  group_by(country_name) %>% 
  summarise(avg_population = mean(population, na.rm = TRUE), .groups = "drop")

# Create `disaster_percap` dataframe.
disaster_percap <- top10_df %>% 
  # Join population average dataframe to disaster totals by country dataframe.
  left_join(pop_avg_top10, by = c("country" = "country_name")) %>% 
  # Calculate the amount of disasters per million people in each country.
  mutate(dis_per_million = country_sum / avg_population * 1e6) %>% 
  # Add flag image source as a new column.
  mutate(iso2 = countrycode(country, origin = "country.name", destination = "iso2c"),
         flag = paste0("https://flagcdn.com/w40/", tolower(iso2), ".png"),
         country = fct_reorder(country, country_sum))

#............................Bar Plot............................

# Create a bar chart.
pop_bar <- ggplot(disaster_percap, aes(x = dis_per_million, y = country)) +
  # Fill by the conditional statement to highlight the Philippines.
  geom_col(aes(fill = country == "Philippines"), alpha = 0.5) +
  # Add the total value to the end of the bar.
  geom_text(
    aes(label = scales::comma(dis_per_million)),
    hjust = -0.85,
    color = pale_blue,
    size = 7,
    family = "sans-serif") +
  # Add flag images to associated country.
  geom_image(aes(image = flag), x = max(disaster_percap$dis_per_million) * 0.03,
             size = 0.05,
             by = "width") +
  # Bars are colored by true/false statements with designated colors.
  scale_fill_manual(
    values = c("TRUE" = "#DFEDF6", "FALSE" = grey_blue),
    guide = "none")  +
  # Expand the x-axis for for flag and total value placement.
  scale_x_continuous(labels = scales::label_comma(),
                   expand = expansion(mult = c(0, 0.18))) +
  labs(x = "Number of Natural Disasters per Million People") +
  theme_minimal(base_size = 17) +
  # Theme adjustments.
  theme(
    axis.text = element_text(family = "sans-serif",
                             size = rel(1.5)),
    axis.title.x = element_text(family = "bricolage",
                                size = rel(2.3), 
                                color = pale_blue),
    axis.title.y = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_text(size = rel(1),
                               margin = margin(r = -0.5),
                               family = "bricolage",
                               color = pale_blue),
    panel.grid = element_blank(),
    panel.grid.major.y = element_blank(),
    legend.position = "none",
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background  = element_rect(fill = "transparent", color = NA))
showtext_opts(dpi = 300)

ggsave(
  filename = here::here("images", "pop_bar.pdf"),
  plot = pop_bar,
  device = cairo_pdf,
  width = 14.5,
  height = 8,
  units = "in")

ggsave(
  filename = here::here("images", "pop_bar.png"),
  plot = pop_bar,
  device = png,
  width = 14.5,
  height = 8,
  units = "in")
```


### Stacked Bar (with country shape)
```{r}
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                 Proportion of Natural Disasters by Country               ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

#.........................Summarize Data.........................
# Create top 5 country list from referencing `top_10` bar plot.
top5_countries <- c("United States", "Philippines", "India", "China", "Japan")

# Create a `type_country` dataframe.
type_country <- nat_dis %>% 
  filter(country %in% top5_countries) %>% 
  # Calculate the proportion of types of disaster for each country.
  group_by(country, disastertype) %>% 
  summarise(disastertype_count = n(), 
            .groups = "drop")

#........................Stacked Bar Plot........................

# Create a stacked bar chart.
stacked_bar <- ggplot(data = type_country, aes(x = country, y = disastertype_count, fill = disastertype, color = "white")) +
  geom_col(position = "fill", alpha = 0.65) +
  # Color disaster type by designated palette.
  scale_fill_manual(values = pal,
                    labels = c("drought" = "Drought",
                               "earthquake" = "Earthquake",
                               "extreme temperature" = "Extreme\nTemperature",
                               "flood" = "Flood",
                               "storm" = "Storm")) +
  scale_color_manual(values = "white")+
  # Make y-axis labels appear as percents.
  scale_y_continuous(labels = scales::label_percent(scale = 100)) +
  labs(x = "Country",
       y = "Percentage of Disasters",
       title = "Distribution of Disaster Types, by Percentages",
       fill = "Disaster Type") +
  theme_minimal() +
  # Theme adjustments.
  theme(
    plot.title = ggtext::element_markdown(family = "bricolage",
                                          face = "bold",
                                          lineheight = 1.2),
    axis.text = element_text(family = "sans-serif",
                             size = rel(0.7)),
    axis.title.x = element_blank(),
    axis.title.y = element_text(family = "bricolage",
                                size = rel(0.75), 
                                margin = margin(t = 15)),
    legend.position = "bottom",
    legend.title.position = "bottom",
    legend.direction = "horizontal",
    legend.text = element_text(family = "bricolage",
                                size = rel(0.75), 
                                margin = margin(t = 2)),
    legend.title = element_blank(),
    legend.key.width = unit(0.5, "cm"),
    legend.key.height = unit(0.5, "cm"),
    panel.grid = element_blank())

ggsave(
  filename = here::here("images", "stacked_bar.pdf"),
  plot = stacked_bar,
  device = cairo_pdf,
  width = 8,
  height = 5,
  units = "in")

ggsave(
  filename = here::here("images", "stacked_bar.png"),
  plot = stacked_bar,
  device = png,
  width = 8,
  height = 5,
  units = "in")
```

```{r}
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##              Population Displaced from Disasters (2008-2018)             ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

#.........................Summarize Data.........................

# Create top 5 country list from referencing `top_10` bar plot.
top5_list <- c("United States", "Philippines", "India", "China", "Japan")

# Create a `displaced_top5` dataframe.
displaced_top5 <- displaced %>%
  # Filter data by the designated time frame (2008-2018) and countries of interest.
  filter(Entity %in% top5_list,
         Year >= 2008, Year <= 2018) %>%
  # Calculate the average amount of people displaced by country.
  group_by(Entity) %>%
  summarise(avg_displaced = mean(displaced, na.rm = TRUE), .groups = "drop")

# Create a `pop_top5` dataframe.
pop_avg_top5 <- pop_avg_top10 %>% 
  # Filter data by countries of interest.
  filter(country_name %in% top5_list)

# Create a `displaced_pop` dataframe.
displaced_pop <- displaced_top5 %>%
  # Join displaced dataframe to population average dataframe.
  left_join(pop_avg_top5, by = c("Entity" = "country_name")) %>%
  # Calculate the amount of people displaced per million people.
  mutate(displaced_per_million = avg_displaced / avg_population * 1e6,
  # Divide by 1000 so that 1 square represents ~1000 displaced people per million.
  # Cap the value at 100 so it fits in a 10x10 waffle grid.
         displaced_sq = pmin(100, round(displaced_per_million / 1000)),
  # Calculate people not displaced.
         not_displaced_sq = 100 - displaced_sq) %>%
  # Select columns needed for waffle plot.
  select(Entity, displaced_sq, not_displaced_sq) %>%
  # Make data long.
  pivot_longer(
    cols = c(displaced_sq, not_displaced_sq),
    names_to = "group",
    values_to = "n") %>%
  # Rename category levels.
  mutate(group = recode(group,
                        displaced_sq = "Displaced",
                        not_displaced_sq = "Not displaced"))

#..........................Waffle Plot...........................

# Create waffle chart.
waffle <- ggplot(displaced_pop, aes(fill = group, values = n)) +
  geom_waffle(n_rows = 10,
              size = 0.1,
              color = "white") +
  # Create a waffle chart for each country.
  facet_wrap(~ Entity, ncol = 5) +
  # Set displacement colors.
  scale_fill_manual(values = c("Displaced" = "navy",
                               "Not displaced" = "grey85")) +
  labs(
    subtitle = "Each 10×10 waffle:\n1 square ≈ 1,000 displaced people per million",
    fill = NULL) +
  theme_minimal() +
  # Theme adjustments.
  theme(
    plot.subtitle = ggtext::element_markdown(family = "bricolage",
                                             lineheight = .5,
                                             size = rel(1),
                                             margin = margin(b = 1),
                                             hjust = 0.5),
    plot.margin = margin(t = 2, r = 2, b = 2, l = 2),
    panel.spacing = unit(0.1, "lines"),
    panel.grid = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    strip.text = element_text(family = "bricolage",
                              size = rel(2),
                              margin = margin(b = .5,
                                              t = 1)),
    legend.position = "none",
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background  = element_rect(fill = "transparent", color = NA))

showtext_opts(dpi = 300)

ggsave(
  filename = here::here("images", "waffle.pdf"),
  plot = waffle,
  device = cairo_pdf,
  width = 2,
  height = 14.5,
  units = "in")

ggsave(
  filename = here::here("images", "waffle.png"),
  plot = waffle,
  device = png,
  width = 14,
  height = 6,
  units = "in")

```

#### Experimenting with other graphs

### Stream Graph
```{r}
#| fig-width: 14
#| fig-height: 12

# # get the top 25 countries
# top25 <- nat_dis %>% 
#   count(country, name = "country_sum") %>% 
#   slice_max(country_sum, n = 10) %>% 
#   pull(country)
# 
# # filter nat_dis based on top 25
# nat_dis %>% 
#   filter(country %in% top25) %>% 
#   count(country, year, disastertype, name = "disaster_count") %>% 
#   # create stream graph
#   ggplot(aes(x = year, y = disaster_count, fill = disastertype)) +
#   geom_stream(type = "proportional") +
#   geom_vline(xintercept = 1980, 
#              col = "grey40") +
#   scale_fill_manual(values = pal,
#                     labels = c("drought" = "Drought",
#                                "earthquake" = "Earthquake",
#                                "extreme temperature" = "Extreme\nTemperature",
#                                "flood" = "Flood",
#                                "storm" = "Storm")) +
#    scale_x_continuous(breaks = c(1960, 1980, 2000, 2018)) +
#   # facet by country
#   facet_wrap(~ country, ncol = 5) +
#   labs(title = "Disaster Type Proportion Between 1980-2018 (Top 25 Countries by Total Disasters)",
#        x = "Year",
#        y = "Proportion of Disasters",
#        fill = "Disaster type") +
#   theme_minimal() +
#   theme(
#     plot.title = ggtext::element_markdown(family = "sarala",
#                                           face = "bold",
#                                           lineheight = 1.2),
#     axis.text = element_text(family = "red",
#                              size = rel(0.7)),
#     axis.title.x = element_blank(),
#     axis.title.y = element_text(family = "noto-sans",
#                                 size = rel(0.75), 
#                                 margin = margin(t = 15,
#                                                 r = 15)),
#     legend.position = "bottom",
#     legend.title.position = "bottom",
#     legend.direction = "horizontal",
#     legend.text = element_text(family = "red",
#                                 size = rel(0.75), 
#                                 margin = margin(t = 2)),
#     legend.title = element_blank(),
#     legend.key.width = unit(0.5, "cm"),
#     panel.spacing.x = unit(1.3, "lines"))
```

# top 10 stream
```{r}
# library(dplyr)
# library(ggplot2)
# library(ggstream)
# 
# top10_stream <- nat_dis %>% 
#   count(country, name = "country_sum") %>% 
#   slice_max(country_sum, n = 10) %>% 
#   pull(country)
# 
# # totals for labeling
# country_totals25 <- nat_dis %>% 
#   filter(country %in% top10_stream) %>% 
#   count(country, name = "country_sum")
# 
# # data for stream
# plot_df <- nat_dis %>% 
#   filter(country %in% top10_stream) %>% 
#   count(country, year, disastertype, name = "disaster_count") %>% 
#   left_join(country_totals25, by = "country") %>% 
#   mutate(country_lab = paste0(country, "\n(n = ", scales::comma(country_sum), ")")) %>% 
#   mutate(
#     country_lab = factor(
#       country_lab,
#       levels = unique(country_lab[order(country_sum, decreasing = TRUE)])
#     ))
# 
# top10_stream <- ggplot(plot_df, aes(x = year, y = disaster_count, fill = disastertype)) +
#   geom_stream(type = "proportional",
#               alpha = 0.8) +
#   geom_vline(xintercept = 1980, 
#              color = "firebrick",
#              alpha = 0.8) +
#   scale_fill_manual(values = pal, labels = c(
#     "drought" = "Drought",
#     "earthquake" = "Earthquake",
#     "extreme temperature" = "Extreme Temperature",
#     "flood" = "Flood",
#     "storm" = "Storm")) +
#   scale_x_continuous(breaks = c(seq(1960, 2018, by = 20), 2018)) +
#   scale_y_continuous(breaks = c(0, 0.5, 1)) +
#   facet_wrap(~ country_lab, ncol = 5) +
#   labs(
#     title = "Disaster Type Proportion Between 1980–2018 (Top 25 Countries by Total Disasters)",
#     x = "Year",
#     y = "Proportion of Disasters",
#     fill = "Disaster type"
#   ) +
#   theme_minimal() +
#   theme(
#     plot.title = ggtext::element_markdown(family = "sarala",
#                                           face = "bold",
#                                           lineheight = 1.2,
#                                           size = rel(0.8),
#                                           hjust = 0.5),
#     axis.text = element_text(family = "red",
#                              size = rel(0.7)),
#     axis.title.x = element_blank(),
#     axis.title.y = element_text(family = "noto-sans",
#                                 size = rel(0.75), 
#                                 margin = margin(r = 10)),
#     legend.position = "bottom",
#     legend.title = element_blank(),
#     legend.key.size = unit(0.5, "cm"),
#     legend.text = element_text(family = "noto-sans",
#                                face = "bold",
#                                size = rel(0.8)),
#     panel.spacing.x = unit(1.3, "lines"),
#     strip.text = element_text(family = "red",
#                               size = rel(0.6)),
#     panel.grid.major = element_line(color = "grey40",
#                                     linewidth = rel(0.3)),
#     panel.grid.minor.x = element_blank(),
#     panel.grid.minor.y = element_blank())
# top10_stream
# 
# showtext_opts(dpi = 300)
# 
# ggsave(
#   filename = here::here("images", "top10_stream.pdf"),
#   plot = top10_stream,
#   device = cairo_pdf,
#   width = 8.5,
#   height = 3,
#   units = "in")
```


```{r}
# dis_counts <- nat_dis %>% 
#   filter(country %in% top25) %>% 
#   count(country, year, disastertype, name = "disaster_count") %>%
#   mutate(country_lab = paste0(country, "\n(n = ", scales::comma(country_sum), ")"))
# 
# ggplot(plot_df,
#        aes(x = year, y = disaster_count, color = disastertype)) +
#   geom_jitter() +
#   geom_vline(xintercept = 1980, color = "grey40") +
#   scale_color_manual(values = pal,
#                      labels = c(
#                        "drought" = "Drought",
#                        "earthquake" = "Earthquake",
#                        "extreme temperature" = "Extreme\nTemperature",
#                        "flood" = "Flood",
#                        "storm" = "Storm"
#                      )) +
#   scale_x_continuous(breaks = c(1960, 1980, 2000, 2018)) +
#   facet_wrap(~ country_lab, ncol = 5, scales = "free_y") +
#   labs(
#     title = "Raw Counts of Natural Disasters by Type (Top 25 Countries)",
#     x = "Year",
#     y = "Number of Disasters",
#     color = "Disaster Type"
#   ) +
#   theme_minimal()
```

```{r}
# displaced <- read_csv(here::here("data", "internally-displaced-persons-from-disasters", "internally-displaced-persons-from-disasters.csv")) %>% 
#   rename("displaced" = "Internally displaced persons, new displacement associated with disasters (number of cases)")
```

```{r}
# top_displaced <- displaced %>% 
#   group_by(Entity) %>% 
#   summarise(total_displaced = sum(displaced,
#                                   na.rm = TRUE)) %>% 
#   slice_max(total_displaced, n = 25) %>% 
#   filter(Entity %in% c(
#     "United States", "Philippines", "India", "China", "Japan",
#     "Thailand", "Indonesia", "Bangladesh", "Vietnam", "Peru"
#   )) %>% 
#   mutate(iso2 = countrycode(Entity, origin = "country.name", destination = "iso2c"),
#          flag = paste0("https://flagcdn.com/w40/", tolower(iso2), ".png"),
#          country = factor(Entity, 
#                           levels = c("United States", "Philippines", "India","China", "Japan", "Thailand", "Indonesia", "Bangladesh", "Vietnam", "Peru"
#   )))
# 
# ggplot(top_displaced,
#        aes(x = total_displaced,
#            y = country)) +
#   geom_col(fill = "grey40") +
#   geom_image(aes(image = flag), x = 0 * max(top10$country_sum), size = rel(0.05)) +
#   scale_x_continuous(labels = scales::label_comma()) +
#   labs(
#     title = "Countries with the Most Disaster Displacement (Internally)",
#     x = "Total People Displaced",
#     y = "Country"
#   ) +
#   theme_minimal()
```




```{r}

# pop_displaced <- displaced %>%
#   left_join(
#     pop_long %>% 
#       select(country_name, year, population),
#     by = c("Entity" = "country_name", "Year" = "year")
#   )
# 
# pop_displaced_top10 <- pop_displaced %>%
#   filter(Entity %in% c("United States", "Philippines", "India","China", "Japan", "Thailand", "Indonesia", "Bangladesh", "Vietnam", "Peru"),
#          Year %in% 2008:2018) %>% 
#   mutate(displaced_per_million = displaced / population * 1e6) %>% 
#   group_by(Entity) %>%
#   summarise(
#     avg_displaced_per_million = mean(displaced_per_million, na.rm = TRUE),
#     .groups = "drop") %>% 
# mutate(iso2 = countrycode(Entity, origin = "country.name", destination = "iso2c"),
#          flag = paste0("https://flagcdn.com/w40/", tolower(iso2), ".png"),
#        Entity = factor(Entity, levels = rev(c("United States", "Philippines", "India","China", "Japan", "Thailand", "Indonesia", "Bangladesh", "Vietnam", "Peru"
#   ))))
# 
# displaced_bar <- ggplot(pop_displaced_top10,
#        aes(x = avg_displaced_per_million, y = Entity)) +
#   geom_col(fill = "black",
#            alpha = 0.5) +
#   geom_image(aes(image = flag), x = 0 * max(top10_df$country_sum), size = rel(0.05)) +
#   labs(
#     title = "Amount of People per Million Internally Displaced by Disasters",
#     x = "Displaced per Million People") +
#   scale_x_continuous(labels = scales::label_comma(),
#                      limits = c(0, 40000),
#                      breaks = c(seq(0, 40000, by = 10000), 40000)) +
#   theme_minimal()+
#   theme(
#     plot.title = ggtext::element_markdown(family = "sarala",
#                                           face = "bold",
#                                           lineheight = 1.2,
#                                           size = rel(0.4),
#                                           hjust = 0.5),
#     axis.text = element_text(family = "red",
#                              size = rel(0.5)),
#     axis.text.y = element_text(size = rel(0.5),
#                                margin = margin(r = -0.5)),
#     axis.text.x = element_text(size = rel(0.5)),
#     axis.title.x = element_text(family = "noto-sans",
#                                 size = rel(0.2)),
#     axis.title.y = element_blank(),
#     panel.grid = element_line(linewidth = 0.1),
#     panel.grid.major.y = element_blank(),
#     plot.caption = ggtext::element_markdown(family = "noto-sans",
#                                             face = "italic",
#                                             color = "grey50",
#                                             size = 1.5,
#                                             margin = margin(t = 0)))
# 
# 
# showtext_opts(dpi = 300)
# 
# ggsave(
#   filename = here::here("images", "displaced.pdf"),
#   plot = displaced_bar,
#   device = cairo_pdf,
#   width = 2.5,
#   height = 1.5,
#   units = "in")
```

```{r}
# library(sf)
# library(ggplot2)
# library(rnaturalearth)
# 
# world <- ne_countries(scale = "medium", returnclass = "sf")
# 
# world <- ggplot() +
#   geom_sf(data = world, fill = "white", color = "black", linewidth = 0.1) +
#   coord_sf(crs = "+proj=robin", expand = FALSE) +
#   theme_void() +
#   theme(
#     panel.grid.major = element_line(color = "grey65", linewidth = 0.1),
#     panel.grid.minor = element_line(color = "grey65", linewidth = 0.1),
#     panel.background = element_rect(fill = "white", color = NA),
#     plot.background = element_rect(fill = "white", color = NA),
#   )
# 
# # ggsave(
# #   "world.png",
# #   plot = world,
# #   width = 3.5,
# #   height = 1.5,
# #   dpi = 300,
# #   bg = "transparent"
# # )
# 
# showtext_opts(dpi = 300)
# 
# ggsave(
#   filename = here::here("images", "world.pdf"),
#   plot = world,
#   device = cairo_pdf,
#   width = 4.5,
#   height = 1.6,
#   units = "in",
#   bg = "transparent")

```
