---
title: "Natural Disasters - Creating Viz"
author: "Jaslyn Miura"
format: pdf
date: 2026-03-08
---

**Goal**: The goal with this data is to explore how the types and amount of natural disasters that have occurred around the world have varied since 1960-2018.

**Questions of Interest:** 
What type natural disasters occur the most around the world?
Which countries experience the most natural disasters per year around the world?
How have the distribution of natural disasters changed over the years (type and count)? Could the changing trends identify major shifts, such as the rise of climate change?

## Introduction
**1. Open with an engaging introduction, which states any necesssary background, your motivation, question(s), data / data source(s), and goal(s) of your visualization.**


### Final Infographic
**2. Embed / render your final infographic following your introduction. If you are embedding a png file, ensure that it’s centered and an apppropriate size. The same applies if you are rendering your infographic from code, though you should additionally ensure that:**

```{r, warning=FALSE, message=FALSE}
# load necessary libraries
library(tidyverse)
library(dplyr)
library(paletteer)
library(lubridate)
library(showtext)
library(sysfonts)
library(stringr)
library(ggstream)
library(cowplot)
library(ggplot2)
library(rnaturalearth)
library(sf)
#install.packages("ggimage")
library(ggimage)
#install.packages("countrycode")
library(countrycode)
library(janitor)
```

## Loading Data
```{r, message=FALSE}
# read in natural disaster world data
nat_dis <- read_csv(here::here("data", "pend-gdis-1960-2018-disasterlocations-csv", "pend-gdis-1960-2018-disasterlocations.csv")) %>%
  mutate(year = as.integer(year)) %>% 
  filter(disastertype %in% c("earthquake", "flood", "storm", "drought", "extreme temperature"))

pop <- read_csv(here::here("data", "world_population", "population.csv"), skip = 4, show_col_types = FALSE) %>% 
  select(where(~ !all(is.na(.))))
```

```{r}
# cleaning population data
pop_long <- pop %>% 
  pivot_longer(
    cols = `1960`:`2024`,
    names_to = "year",
    values_to = "population") %>% 
  mutate(year = as.integer(year)) %>% 
  clean_names() %>% 
  select(country_name, year, population, country_code) %>% 
  mutate(country_name = recode(country_name,
                               "United States of America" = "United States",
                               "Viet Nam" = "Vietnam"))
```


```{r}
# add special fonts
font_add_google(name = "Sarala", family = "sarala")
font_add_google(name = "Noto Sans", family = "noto-sans")
font_add_google(name = "Red Hat Mono", family = "red")
```


### Line Graph
```{r, message=FALSE}

showtext.auto(enable = TRUE)

# summarize the count of natural disasters
disaster_counts_line <- nat_dis %>% 
  count(year, name = "total_disasters")

# create a line graph
line_trend <- ggplot(data = disaster_counts_line, aes(x = year, y = total_disasters)) +
  # add a vline to represent climate change
  geom_vline(xintercept = 1980,
             color = "firebrick",
             alpha = 0.7) +
  geom_line() +
  geom_point(size = 0.4) +
  scale_x_continuous(limits = c(1960, 2018),
                     breaks = c(seq(1960, 2015, by = 20), 2018)) +
  scale_y_continuous(labels = scales::label_comma()) +
  labs(y = "Total Natural Disasters",
       caption = "Data Source: NASA EarthData") +
  theme_minimal() +
  annotate("text",
           x = 1980,
           y = 1000,
           label = "Global Warming\nAcceleration",
           angle = 90,
           vjust = -0.5,
           family = "red",
           size = rel(1.6),
           color = "firebrick") +
  theme(
    plot.caption = ggtext::element_markdown(family = "noto-sans",
                                            face = "italic",
                                            color = "grey50",
                                            size = rel(0.3),
                                            margin = margin(t = 3)),
    axis.text = element_text(family = "red"),
    axis.text.y = element_text(size = rel(0.7)),
    axis.text.x = element_text(size = rel(0.8)),
    axis.title.x = element_blank(),
    axis.title.y = element_text(family = "noto-sans",
                                size = rel(0.5), 
                                margin = margin(t = 15,
                                                r = 12)),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background  = element_rect(fill = "transparent", color = NA))

#...................set resolution to match GD...................
showtext_opts(dpi = 300)

ggsave(
  filename = here::here("images", "line_trend.pdf"),
  plot = line_trend,
  device = cairo_pdf,
  width = 7, 
  height = 2.5,
  units = "in")

ggsave(
  filename = here::here("images", "line_trend.png"),
  plot = line_trend,
  device = png,
  width = 7, 
  height = 2.5,
  units = "in")

```

### Bar Graph with Flags -- **FINAL**
```{r}
showtext.auto(enable = TRUE)

# summarize the top 10 countries
top10_df <- nat_dis %>% 
  count(country, year, name = "disaster_count") %>% 
  group_by(country) %>% 
  summarize(country_sum = sum(disaster_count), 
            .groups = "drop") %>%  
  slice_max(country_sum, n = 10) %>%             
  mutate(iso2 = countrycode(country, origin = "country.name", destination = "iso2c"),
         flag = paste0("https://flagcdn.com/w40/", tolower(iso2), ".png"),
         country = fct_reorder(country, country_sum))

# create a bar chart
top10 <- ggplot(top10_df, aes(x = country_sum, y = country)) +
  geom_col(fill = "#B8B8B8",
           alpha = 0.5) +
  geom_text(aes(label = scales::comma(country_sum)),
            hjust = -0.4, 
            color = "black",
            size = rel(1),
            family = "red") +
  # add flag images to associated country
  geom_image(aes(image = flag),
             x = max(top10_df$country_sum) * 0.03,
             size = 0.05,
             by = "width") +
scale_x_continuous(labels = scales::label_comma(),
                   expand = expansion(mult = c(0, 0.1))) +
  labs(title = "Countries with the Most Recorded Natural Disasters",
       subtitle = "From 1960-2018, the United States and the Philippines recorded the highest total number of disasters.",
       x = "Number of Natural Disasters",
       y = "Country",
       caption = "Data Source: NASA EarthData") +
  theme_minimal() +
  theme(
    plot.title = ggtext::element_markdown(family = "sarala",
                                          face = "bold",
                                          size = rel(0.5),
                                          margin = margin(b = rel(0.8))),
    plot.subtitle = ggtext::element_markdown(family = "sarala",
                                             size = rel(0.2),
                                             margin = margin(b = rel(0.8))),
    axis.text = element_text(family = "red",
                             size = rel(0.5)),
    axis.text.y = element_text(size = rel(0.5),
                               color = "black"),
    axis.text.x = element_text(size = rel(0.5),
                               color = "black"),
    axis.title.x = element_text(family = "noto-sans",
                                size = rel(0.2),
                                color = "black"),
    axis.title.y = element_blank(),
    panel.grid = element_line(linewidth = 0.1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    plot.caption = ggtext::element_markdown(family = "noto-sans",
                                            face = "italic",
                                            color = "grey50",
                                            size = 1.5,
                                            margin = margin(t = 0)),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background  = element_rect(fill = "transparent", color = NA))
top10

showtext_opts(dpi = 300)

ggsave(
  filename = here::here("images", "top10.pdf"),
  plot = top10,
  device = cairo_pdf,
  width = 2.5, 
  height = 2,
  units = "in")

ggsave(
  filename = here::here("images", "top10.png"),
  plot = top10,
  device = png,
  width = 2.5, 
  height = 2,
  units = "in")


```


### Population Bar Chart
```{r}

library(gghighlight)

pop_avg_top10 <- pop_long %>% 
  filter(year >= 1960, year <= 2018) %>% 
  group_by(country_name) %>% 
  summarise(avg_population = mean(population, na.rm = TRUE), .groups = "drop") %>% 
  filter(country_name %in% c(
    "United States","Philippines","India","China","Japan",
    "Thailand","Indonesia","Bangladesh","Vietnam","Peru"
  ))

#mismatched <- anti_join(country_totals, pop_long, by = c("name" = "country_name"))

top10_percap <- top10_df %>% 
  left_join(pop_avg_top10, by = c("country" = "country_name")) %>% 
  mutate(dis_per_million = country_sum / avg_population * 1e6) %>% 
  mutate(iso2 = countrycode(country, origin = "country.name", destination = "iso2c"),
         flag = paste0("https://flagcdn.com/w40/", tolower(iso2), ".png"),
         country = fct_reorder(country, country_sum))

pop_bar <- ggplot(top10_percap, aes(x = dis_per_million, y = country)) +
  geom_col(aes(fill = country == "Philippines"), alpha = 0.9) +
  geom_text(
    aes(label = scales::comma(dis_per_million)),
    hjust = -0.55,
    color = "black",
    size = rel(1),
    family = "red"
  ) +
  # add flag images to associated country
  geom_image(aes(image = flag), x = max(top10_percap$dis_per_million) * 0.035,
             size = 0.05,
             by = "width") +
  scale_fill_manual(
    values = c("TRUE" = "#000080", "FALSE" = "#B8B8B8"),
    guide = "none"
  )  +
scale_x_continuous(labels = scales::label_comma(),
                   expand = expansion(mult = c(0, 0.18))) +
  labs(title = "Disaster Exposure per Million People",
       subtitle = "When population is considered, smaller countries like the Philippines experience far more disasters relative to population size.",
       x = "Number of Natural Disasters per Million People",
       caption = "Data Source: NASA EarthData and World Bank Group") +
  theme_minimal() +
  theme(
    plot.title = ggtext::element_markdown(family = "sarala",
                                          face = "bold",
                                          size = rel(0.5),
                                          margin = margin(b = rel(0.8))),
    plot.subtitle = ggtext::element_markdown(family = "sarala",
                                             size = rel(0.2),
                                             margin = margin(b = rel(0.8))),
    plot.caption = ggtext::element_markdown(family = "noto-sans",
                                            face = "italic",
                                            color = "grey50",
                                            size = 1.5,
                                            margin = margin(t = 0)),
    axis.text = element_text(family = "red",
                             size = rel(0.5)),
    axis.title.x = element_text(family = "noto-sans",
                                size = rel(0.2)),
    axis.title.y = element_blank(),
    axis.text.x = element_text(size = rel(0.5)),
    axis.text.y = element_text(size = rel(0.5),
                               margin = margin(r = -0.5)),
    panel.grid = element_line(linewidth = 0.1),
    panel.grid.major.y = element_blank(),
    legend.position = "none",
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background  = element_rect(fill = "transparent", color = NA))
pop_bar


showtext_opts(dpi = 300)

ggsave(
  filename = here::here("images", "pop_bar.pdf"),
  plot = pop_bar,
  device = cairo_pdf,
  width = 2.5, 
  height = 2,
  units = "in")

ggsave(
  filename = here::here("images", "pop_bar.png"),
  plot = pop_bar,
  device = png,
  width = 2.5, 
  height = 2,
  units = "in")
```


### Stacked Bar (with country shape)
```{r}

# create a color palette for disasters
pal <- c("drought" = "#9D836C",
         "earthquake" = "#70815F",
         "extreme temperature" = "#805D93",
         "flood" = "#7B9EA8",
         "storm" = "#F3DFA2")

# summarize data by country and type
stacked_df <- nat_dis %>% 
  filter(country %in% c("United States", "Philippines", "India", "China", "Japan")) %>% 
  group_by(country, disastertype) %>% 
  summarise(disastertype_count = n(), 
            .groups = "drop") %>% 
  mutate(country = fct_reorder(country, disastertype_count)) 
# create a stacked bar chart
stacked_bar <- ggplot(data = stacked_df, aes(x = country, y = disastertype_count, fill = disastertype)) +
  geom_col(position = "fill", alpha = 0.7) +
  scale_fill_manual(values = pal,
                    labels = c("drought" = "Drought",
                               "earthquake" = "Earthquake",
                               "extreme temperature" = "Extreme\nTemperature",
                               "flood" = "Flood",
                               "storm" = "Storm")) +
  scale_y_continuous(labels = scales::label_percent(scale = 100)) +
  labs(x = "Country",
       y = "Percentage of Disasters",
       title = "Distribution of Disaster Types, by Percentages",
       fill = "Disaster Type") +
  theme_minimal() +
  theme(
    plot.title = ggtext::element_markdown(family = "sarala",
                                          face = "bold",
                                          lineheight = 1.2),
    axis.text = element_text(family = "red",
                             size = rel(0.7)),
    axis.title.x = element_blank(),
    axis.title.y = element_text(family = "noto-sans",
                                size = rel(0.75), 
                                margin = margin(t = 15)),
    legend.position = "bottom",
    legend.title.position = "bottom",
    legend.direction = "horizontal",
    legend.text = element_text(family = "red",
                                size = rel(0.75), 
                                margin = margin(t = 2)),
    legend.title = element_blank(),
    legend.key.width = unit(0.5, "cm"),
    legend.key.height = unit(0.5, "cm"))

ggsave(
  filename = here::here("images", "stacked_bar.png"),
  plot = stacked_bar,
  device = png,
  width = 3.5,
  height = 1.6,
  units = "in")

```


### Stream Graph
```{r}
#| fig-width: 14
#| fig-height: 12

# # get the top 25 countries
# top25 <- nat_dis %>% 
#   count(country, name = "country_sum") %>% 
#   slice_max(country_sum, n = 10) %>% 
#   pull(country)
# 
# # filter nat_dis based on top 25
# nat_dis %>% 
#   filter(country %in% top25) %>% 
#   count(country, year, disastertype, name = "disaster_count") %>% 
#   # create stream graph
#   ggplot(aes(x = year, y = disaster_count, fill = disastertype)) +
#   geom_stream(type = "proportional") +
#   geom_vline(xintercept = 1980, 
#              col = "grey40") +
#   scale_fill_manual(values = pal,
#                     labels = c("drought" = "Drought",
#                                "earthquake" = "Earthquake",
#                                "extreme temperature" = "Extreme\nTemperature",
#                                "flood" = "Flood",
#                                "storm" = "Storm")) +
#    scale_x_continuous(breaks = c(1960, 1980, 2000, 2018)) +
#   # facet by country
#   facet_wrap(~ country, ncol = 5) +
#   labs(title = "Disaster Type Proportion Between 1980-2018 (Top 25 Countries by Total Disasters)",
#        x = "Year",
#        y = "Proportion of Disasters",
#        fill = "Disaster type") +
#   theme_minimal() +
#   theme(
#     plot.title = ggtext::element_markdown(family = "sarala",
#                                           face = "bold",
#                                           lineheight = 1.2),
#     axis.text = element_text(family = "red",
#                              size = rel(0.7)),
#     axis.title.x = element_blank(),
#     axis.title.y = element_text(family = "noto-sans",
#                                 size = rel(0.75), 
#                                 margin = margin(t = 15,
#                                                 r = 15)),
#     legend.position = "bottom",
#     legend.title.position = "bottom",
#     legend.direction = "horizontal",
#     legend.text = element_text(family = "red",
#                                 size = rel(0.75), 
#                                 margin = margin(t = 2)),
#     legend.title = element_blank(),
#     legend.key.width = unit(0.5, "cm"),
#     panel.spacing.x = unit(1.3, "lines"))
```
# top 10 stream
```{r}
# library(dplyr)
# library(ggplot2)
# library(ggstream)
# 
# top10_stream <- nat_dis %>% 
#   count(country, name = "country_sum") %>% 
#   slice_max(country_sum, n = 10) %>% 
#   pull(country)
# 
# # totals for labeling
# country_totals25 <- nat_dis %>% 
#   filter(country %in% top10_stream) %>% 
#   count(country, name = "country_sum")
# 
# # data for stream
# plot_df <- nat_dis %>% 
#   filter(country %in% top10_stream) %>% 
#   count(country, year, disastertype, name = "disaster_count") %>% 
#   left_join(country_totals25, by = "country") %>% 
#   mutate(country_lab = paste0(country, "\n(n = ", scales::comma(country_sum), ")")) %>% 
#   mutate(
#     country_lab = factor(
#       country_lab,
#       levels = unique(country_lab[order(country_sum, decreasing = TRUE)])
#     ))
# 
# top10_stream <- ggplot(plot_df, aes(x = year, y = disaster_count, fill = disastertype)) +
#   geom_stream(type = "proportional",
#               alpha = 0.8) +
#   geom_vline(xintercept = 1980, 
#              color = "firebrick",
#              alpha = 0.8) +
#   scale_fill_manual(values = pal, labels = c(
#     "drought" = "Drought",
#     "earthquake" = "Earthquake",
#     "extreme temperature" = "Extreme Temperature",
#     "flood" = "Flood",
#     "storm" = "Storm")) +
#   scale_x_continuous(breaks = c(seq(1960, 2018, by = 20), 2018)) +
#   scale_y_continuous(breaks = c(0, 0.5, 1)) +
#   facet_wrap(~ country_lab, ncol = 5) +
#   labs(
#     title = "Disaster Type Proportion Between 1980–2018 (Top 25 Countries by Total Disasters)",
#     x = "Year",
#     y = "Proportion of Disasters",
#     fill = "Disaster type"
#   ) +
#   theme_minimal() +
#   theme(
#     plot.title = ggtext::element_markdown(family = "sarala",
#                                           face = "bold",
#                                           lineheight = 1.2,
#                                           size = rel(0.8),
#                                           hjust = 0.5),
#     axis.text = element_text(family = "red",
#                              size = rel(0.7)),
#     axis.title.x = element_blank(),
#     axis.title.y = element_text(family = "noto-sans",
#                                 size = rel(0.75), 
#                                 margin = margin(r = 10)),
#     legend.position = "bottom",
#     legend.title = element_blank(),
#     legend.key.size = unit(0.5, "cm"),
#     legend.text = element_text(family = "noto-sans",
#                                face = "bold",
#                                size = rel(0.8)),
#     panel.spacing.x = unit(1.3, "lines"),
#     strip.text = element_text(family = "red",
#                               size = rel(0.6)),
#     panel.grid.major = element_line(color = "grey40",
#                                     linewidth = rel(0.3)),
#     panel.grid.minor.x = element_blank(),
#     panel.grid.minor.y = element_blank())
# top10_stream
# 
# showtext_opts(dpi = 300)
# 
# ggsave(
#   filename = here::here("images", "top10_stream.pdf"),
#   plot = top10_stream,
#   device = cairo_pdf,
#   width = 8.5,
#   height = 3,
#   units = "in")
```


```{r}
# dis_counts <- nat_dis %>% 
#   filter(country %in% top25) %>% 
#   count(country, year, disastertype, name = "disaster_count") %>%
#   mutate(country_lab = paste0(country, "\n(n = ", scales::comma(country_sum), ")"))
# 
# ggplot(plot_df,
#        aes(x = year, y = disaster_count, color = disastertype)) +
#   geom_jitter() +
#   geom_vline(xintercept = 1980, color = "grey40") +
#   scale_color_manual(values = pal,
#                      labels = c(
#                        "drought" = "Drought",
#                        "earthquake" = "Earthquake",
#                        "extreme temperature" = "Extreme\nTemperature",
#                        "flood" = "Flood",
#                        "storm" = "Storm"
#                      )) +
#   scale_x_continuous(breaks = c(1960, 1980, 2000, 2018)) +
#   facet_wrap(~ country_lab, ncol = 5, scales = "free_y") +
#   labs(
#     title = "Raw Counts of Natural Disasters by Type (Top 25 Countries)",
#     x = "Year",
#     y = "Number of Disasters",
#     color = "Disaster Type"
#   ) +
#   theme_minimal()
```

```{r}
displaced <- read_csv(here::here("data", "internally-displaced-persons-from-disasters", "internally-displaced-persons-from-disasters.csv")) %>% 
  rename("displaced" = "Internally displaced persons, new displacement associated with disasters (number of cases)")
```
```{r}
top_displaced <- displaced %>% 
  group_by(Entity) %>% 
  summarise(total_displaced = sum(displaced,
                                  na.rm = TRUE)) %>% 
  slice_max(total_displaced, n = 25) %>% 
  filter(Entity %in% c(
    "United States", "Philippines", "India", "China", "Japan",
    "Thailand", "Indonesia", "Bangladesh", "Vietnam", "Peru"
  )) %>% 
  mutate(iso2 = countrycode(Entity, origin = "country.name", destination = "iso2c"),
         flag = paste0("https://flagcdn.com/w40/", tolower(iso2), ".png"),
         country = factor(Entity, 
                          levels = c("United States", "Philippines", "India","China", "Japan", "Thailand", "Indonesia", "Bangladesh", "Vietnam", "Peru"
  )))

ggplot(top_displaced,
       aes(x = total_displaced,
           y = country)) +
  geom_col(fill = "grey40") +
  geom_image(aes(image = flag), x = 0 * max(top10$country_sum), size = rel(0.05)) +
  scale_x_continuous(labels = scales::label_comma()) +
  labs(
    title = "Countries with the Most Disaster Displacement (Internally)",
    x = "Total People Displaced",
    y = "Country"
  ) +
  theme_minimal()
```




```{r}

pop_displaced <- displaced %>%
  left_join(
    pop_long %>% 
      select(country_name, year, population),
    by = c("Entity" = "country_name", "Year" = "year")
  )

pop_displaced_top10 <- pop_displaced %>%
  filter(Entity %in% c("United States", "Philippines", "India","China", "Japan", "Thailand", "Indonesia", "Bangladesh", "Vietnam", "Peru"),
         Year %in% 2008:2018) %>% 
  mutate(displaced_per_million = displaced / population * 1e6) %>% 
  group_by(Entity) %>%
  summarise(
    avg_displaced_per_million = mean(displaced_per_million, na.rm = TRUE),
    .groups = "drop") %>% 
mutate(iso2 = countrycode(Entity, origin = "country.name", destination = "iso2c"),
         flag = paste0("https://flagcdn.com/w40/", tolower(iso2), ".png"),
       Entity = factor(Entity, levels = rev(c("United States", "Philippines", "India","China", "Japan", "Thailand", "Indonesia", "Bangladesh", "Vietnam", "Peru"
  ))))

displaced_bar <- ggplot(pop_displaced_top10,
       aes(x = avg_displaced_per_million, y = Entity)) +
  geom_col(fill = "black",
           alpha = 0.5) +
  geom_image(aes(image = flag), x = 0 * max(top10_df$country_sum), size = rel(0.05)) +
  labs(
    title = "Amount of People per Million Internally Displaced by Disasters",
    x = "Displaced per Million People") +
  scale_x_continuous(labels = scales::label_comma(),
                     limits = c(0, 40000),
                     breaks = c(seq(0, 40000, by = 10000), 40000)) +
  theme_minimal()+
  theme(
    plot.title = ggtext::element_markdown(family = "sarala",
                                          face = "bold",
                                          lineheight = 1.2,
                                          size = rel(0.4),
                                          hjust = 0.5),
    axis.text = element_text(family = "red",
                             size = rel(0.5)),
    axis.text.y = element_text(size = rel(0.5),
                               margin = margin(r = -0.5)),
    axis.text.x = element_text(size = rel(0.5)),
    axis.title.x = element_text(family = "noto-sans",
                                size = rel(0.2)),
    axis.title.y = element_blank(),
    panel.grid = element_line(linewidth = 0.1),
    panel.grid.major.y = element_blank(),
    plot.caption = ggtext::element_markdown(family = "noto-sans",
                                            face = "italic",
                                            color = "grey50",
                                            size = 1.5,
                                            margin = margin(t = 0)))


showtext_opts(dpi = 300)

ggsave(
  filename = here::here("images", "displaced.pdf"),
  plot = displaced_bar,
  device = cairo_pdf,
  width = 2.5,
  height = 1.5,
  units = "in")
```


```{r}
library(dplyr)
library(waffle)
library(tidyr)
library(ggplot2)

top5_countries <- c("United States", "Philippines", "India", "China", "Japan")
```


```{r}
pop_top10 <- pop_long %>%
  filter(
    country_name %in% top5_countries,
    year >= 2008, year <= 2018
  ) %>%
  group_by(country_name) %>%
  summarise(avg_population = mean(population, na.rm = TRUE), .groups = "drop")

displaced_top10 <- displaced %>%
  filter(
    Entity %in% top5_countries,
    Year >= 2008, Year <= 2018
  ) %>%
  group_by(Entity) %>%
  summarise(avg_displaced = mean(displaced, na.rm = TRUE), .groups = "drop")

waffle_df <- displaced_top10 %>%
  left_join(pop_top10, by = c("Entity" = "country_name")) %>%
  mutate(
    displaced_per_million = avg_displaced / avg_population * 1e6,
    displaced_sq = pmin(100, round(displaced_per_million / 1000)),
    not_displaced_sq = 100 - displaced_sq
  )

waffle_long <- waffle_df %>%
  select(Entity, displaced_sq, not_displaced_sq) %>%
  pivot_longer(
    cols = c(displaced_sq, not_displaced_sq),
    names_to = "group",
    values_to = "n"
  ) %>%
  mutate(
    group = recode(
      group,
      displaced_sq = "Displaced",
      not_displaced_sq = "Not displaced"
    )
  ) %>% 
  mutate(
    Entity = factor(
      Entity,
      levels = c("United States", "Philippines", "India", "China", "Japan")
    )
  )

waffle <- ggplot(waffle_long, aes(fill = group, values = n)) +
  geom_waffle(n_rows = 10,
              size = 0.25,
              color = "white") +
  facet_wrap(~ Entity, ncol = 5) +
  scale_fill_manual(values = c("Displaced" = "navy",
                               "Not displaced" = "grey85")) +
  labs(
    title = "Average Annual Disaster Displacement<br>Relative to Population",
    subtitle = "Each 10×10 waffle:\n1 square ≈ 1,000 displaced people per million",
    fill = NULL) +
  theme_minimal() +
  theme(
    plot.title = ggtext::element_markdown(family = "sarala",
                                          face = "bold",
                                          size = rel(0.8),
                                          hjust = 0.5,
                                          margin = margin(b = 0.5)),
    plot.subtitle = ggtext::element_markdown(family = "sarala",
                                             lineheight = .5,
                                             size = rel(0.3),
                                             margin = margin(b = 1),
                                             hjust = 0.5),
    plot.margin = margin(t = 5, r = 3, b = 3, l = 5),
    panel.spacing = unit(0.1, "lines"),
    panel.grid = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    strip.text = element_text(family = "red",
                              size = rel(0.4),
                              margin = margin(b = .5,
                                              t = 1)),
    legend.position = "none")


showtext_opts(dpi = 300)

ggsave(
  filename = here::here("images", "waffle.pdf"),
  plot = waffle,
  device = cairo_pdf,
  width = 3.5,
  height = 1.6,
  units = "in")

ggsave(
  filename = here::here("images", "waffle.png"),
  plot = waffle,
  device = png,
  width = 3.5,
  height = 1.6,
  units = "in")

```


```{r}
library(sf)
library(ggplot2)
library(rnaturalearth)

world <- ne_countries(scale = "medium", returnclass = "sf")

world <- ggplot() +
  geom_sf(data = world, fill = "white", color = "black", linewidth = 0.1) +
  coord_sf(crs = "+proj=robin", expand = FALSE) +
  theme_void() +
  theme(
    panel.grid.major = element_line(color = "grey65", linewidth = 0.1),
    panel.grid.minor = element_line(color = "grey65", linewidth = 0.1),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
  )

# ggsave(
#   "world.png",
#   plot = world,
#   width = 3.5,
#   height = 1.5,
#   dpi = 300,
#   bg = "transparent"
# )

showtext_opts(dpi = 300)

ggsave(
  filename = here::here("images", "world.pdf"),
  plot = world,
  device = cairo_pdf,
  width = 4.5,
  height = 1.6,
  units = "in",
  bg = "transparent")

```


```{r}
#install.packages("ggpattern")
library(ggpattern)

library(dplyr)
library(ggplot2)
library(ggpattern)
library(forcats)
library(ggtext)
library(scales)

nat_dis %>% 
  filter(country %in% c("United States", "Philippines", "India", "China", "Japan")) %>% 
  group_by(country, disastertype) %>% 
  summarise(
    disastertype_count = n(),
    .groups = "drop"
  ) %>% 
  mutate(
    country = factor(
      country,
      levels = c("United States", "Philippines", "India", "China", "Japan")
    )
  ) %>% 
  ggplot(
    aes(
      x = country,
      y = disastertype_count,
      fill = disastertype,
      pattern = disastertype
    )
  ) +
  geom_col_pattern(
    position = "fill",
    alpha = 0.9,
    color = "white",
    linewidth = 0.3,
    pattern_fill = "black",
    pattern_colour = "black",
    pattern_alpha = 0.18,
    pattern_density = 0.25,
    pattern_spacing = 0.04
  ) +
  scale_fill_manual(
    values = pal,
    labels = c(
      "drought" = "Drought",
      "earthquake" = "Earthquake",
      "extreme temperature" = "Extreme\nTemperature",
      "flood" = "Flood",
      "storm" = "Storm"
    )
  ) +
  scale_pattern_manual(
    values = c(
      "drought" = "none",
      "earthquake" = "circle",
      "extreme temperature" = "stripe",
      "flood" = "wave",
      "storm" = "crosshatch"
    ),
    labels = c(
      "drought" = "Drought",
      "earthquake" = "Earthquake",
      "extreme temperature" = "Extreme\nTemperature",
      "flood" = "Flood",
      "storm" = "Storm"
    )
  ) +
  scale_y_continuous(labels = label_percent(scale = 100)) +
  labs(
    title = "Distribution of Disaster Types by Country",
    subtitle = "Floods and storms dominate, but the composition of disaster types differs across countries.",
    x = NULL,
    y = "Percentage of Disasters",
    fill = NULL,
    pattern = NULL
  ) +
  theme_minimal() +
  theme(
    plot.title = ggtext::element_markdown(
      family = "sarala",
      face = "bold",
      lineheight = 1.2,
      size = rel(1)
    ),
    plot.subtitle = ggtext::element_markdown(
      family = "noto-sans",
      size = rel(0.8),
      margin = margin(b = 10)
    ),
    axis.text = element_text(
      family = "red",
      size = rel(0.75)
    ),
    axis.title.y = element_text(
      family = "noto-sans",
      size = rel(0.8),
      margin = margin(r = 15)
    ),
    legend.position = "bottom",
    legend.text = element_text(
      family = "red",
      size = rel(0.75)
    ),
    legend.key.width = unit(0.55, "cm"),
    legend.key.height = unit(0.55, "cm"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank()
  ) +
  guides(fill = "none")
```
```{r}
library(dplyr)
library(ggplot2)
library(ggpattern)
library(ggtext)
library(scales)

stacked <- nat_dis %>% 
  filter(country %in% c("United States", "Philippines", "India", "China", "Japan")) %>% 
  group_by(country, disastertype) %>% 
  summarise(
    disastertype_count = n(),
    .groups = "drop"
  ) %>% 
  mutate(
    country = factor(
      country,
      levels = c("United States", "Philippines", "India", "China", "Japan")
    )
  ) %>% 
  group_by(country) %>% 
  mutate(prop = disastertype_count / sum(disastertype_count)) 

  stacked_bar <- ggplot(data = stacked,
    aes(
      x = country,
      y = disastertype_count,
      fill = disastertype
    )
  ) +
  geom_col_pattern(
    position = "fill",
    pattern = "circle",
    alpha = 0.6,
    color = "white",
    linewidth = 0.3,
    pattern_fill = "black",
    pattern_colour = "black",
    pattern_alpha = 0.1,
    pattern_density = 0.08,
    pattern_spacing = 0.08,
    pattern_size = 0.1
  ) +
  scale_fill_manual(
    values = pal,
    labels = c(
      "drought" = "Drought",
      "earthquake" = "Earthquake",
      "extreme temperature" = "Extreme\nTemperature",
      "flood" = "Flood",
      "storm" = "Storm"
    )
  ) +
  scale_y_continuous(labels = label_percent(scale = 100)) +
  labs(
    title = "Distribution of Disaster Types by Country",
    subtitle = "Disaster composition differs across countries, with floods and storms making up the largest shares.",
    x = NULL,
    y = "Percentage of Disasters",
    fill = NULL
  ) +
  theme_minimal() +
  theme(
    plot.title = ggtext::element_markdown(
      family = "sarala",
      face = "bold",
      lineheight = 1.2,
      size = rel(1)
    ),
    plot.subtitle = ggtext::element_markdown(
      family = "noto-sans",
      size = rel(0.8),
      margin = margin(b = 10)
    ),
    axis.text = element_text(
      family = "red",
      size = rel(0.75)
    ),
    axis.title.y = element_text(
      family = "noto-sans",
      size = rel(0.8),
      margin = margin(r = 15)
    ),
    legend.position = "bottom",
    legend.text = element_text(
      family = "red",
      size = rel(0.75)
    ),
    legend.key.width = unit(0.55, "cm"),
    legend.key.height = unit(0.55, "cm"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

ggsave(
  filename = here::here("images", "stacked_bar.pdf"),
  plot = stacked_bar,
  device = cairo_pdf,
  width = 8,
  height = 6,
  units = "in",
  bg = "transparent")
```

```{r}
plot_df <- nat_dis %>% 
  filter(country %in% c("United States", "Philippines", "India", "China", "Japan")) %>% 
  group_by(country, disastertype) %>% 
  summarise(disastertype_count = n(), .groups = "drop") %>% 
  group_by(country) %>% 
  mutate(prop = disastertype_count / sum(disastertype_count))

ggplot(plot_df,
  aes(
    x = country,
    y = disastertype_count,
    fill = disastertype
  )
) +
geom_col_pattern(
  position = "fill",
  pattern = "circle",
  pattern_density = 0.08,
  pattern_spacing = 0.08,
  pattern_size = 0.15,
  pattern_fill = "black",
  pattern_alpha = 0.18,
  color = "white",
  linewidth = 0.3
) +
geom_text(
  aes(label = scales::percent(prop, accuracy = 1)),
  position = position_fill(vjust = 0.5),
  size = 3,
  family = "red",
  color = "white"
)
```



### Body
3. The body of your blog post should address your approach and decisions for the ten design elements listed in the final project description (though you are welcome and encouraged to comment on any others that are not explicitly listed). You may choose to render and refer to individual / component data visualizations throughout the body of your blog post, as necessary or appropriate – if you do so, be sure to suppress code from printing.


